---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---

```{r}
#load libraries
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(tidyr)
library(kableExtra)
library(knitr)
library(fastDummies)
```

```{r}
#Loading original datast
tsa_claims <- read_csv("tsa_claims.csv")

#exploring data summary,  structure and types
str(tsa_claims)
summary(tsa_claims)

```
```{r}
#Renaming the non-standard variables with standard names

TSA_claims <- tsa_claims %>%
  rename(Claim_Number =  `Claim Number`,
         Date_Received = `Date Received`,
         Incident_Date = `Incident Date`,
         Airport_Code = `Airport Code`,
         Airport_Name = `Airport Name`,
         Airline_Name = `Airline Name`,
         Claim_Type = `Claim Type`,
         Claim_Site = `Claim Site`,
         Claim_Amount = `Claim Amount`,
         Close_Amount = `Close Amount`)
TSA_claims
```


```{r}
#Remove unwanted characters from specific variables, specifically Claim_Amount & Close_Amount

TSA_claims <- TSA_claims %>%
  mutate_if(is.character, str_replace_all, pattern = '[$]', replacement = '') %>%
  mutate_if(is.character, str_replace_all, pattern = '[;]', replacement = '') %>%
  #Convert data type for specific variables
  mutate(
         Date_Received = dmy(Date_Received),
         Incident_Date = mdy(Incident_Date),
         Airport_Code = as_factor(Airport_Code),
         Claim_Type = as_factor(Claim_Type),
         Claim_Site = as_factor(Claim_Site),
         Claim_Amount = as.double(Claim_Amount),
         Status = as_factor(Status),
         Close_Amount = as.double(Close_Amount),
         Disposition = as_factor(Disposition)
         )

#Verifying conversion to TSA_claims tbl
str(TSA_claims)
summary(TSA_claims)
```
Investigationg NAs introduced by coersion
```{r}
TSA_claims %>%
  select(Claim_Amount) %>%
  filter(is.na(as.double(Claim_Amount)))
```
```{r}
 airports <- TSA_claims %>%
  select(Claim_Number, Airport_Code, Airport_Name) %>%
  dummy_cols(select_columns = "Airport_Code",
                                   ignore_na = FALSE)

airports

mice::md.pattern(airports)
```


Incident Distribution by year
```{r}
Incident_distibution <- TSA_claims %>%
  mutate(Incident_Year = as.numeric(format(Incident_Date, "%Y")),
         Claim_Year = as.numeric(format(Date_Received, "%Y"))) %>%
  select(Incident_Date, Incident_Year, Date_Received, Claim_Year, Claim_Amount, Close_Amount, Disposition) %>%
  group_by(Incident_Year, Claim_Year) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = Incident_Year),binwidth = 50, stat = "count", color = "slateblue", fill = "slategray") +
  geom_histogram(mapping = aes(x = Claim_Year),binwidth = 50, stat = "count", color = "blue", fill = "gray", alpha = 0.5, fill = "dodge")
  geom_line(aes(x = Year, y = count(Year), color = Disposition)) 

Incident_distibution
  
```



```{r}
options(scipen = 999)

TSA_claims %>%
  select(Close_Amount, Disposition) %>%
  group_by(Disposition) %>%
  arrange(Disposition) %>%
  filter(!is.na(Close_Amount), 
         !is.na(Disposition)) %>%
  mutate(log_CloseAmount = log(Close_Amount)) %>%
  ggplot() +
  geom_histogram(aes(x = Disposition, y= log_CloseAmount, color = Disposition), stat = "identity")
  


TSA_claims %>%
  select(Claim_Amount, Disposition) %>%
  group_by(Disposition) %>%
  arrange(Disposition) %>%
  filter(!is.na(Claim_Amount), 
         !is.na(Disposition)) %>%
  mutate(log_ClaimAmount = log(Claim_Amount)) %>%
  ggplot() +
  geom_histogram(aes(x = Disposition, y= log_ClaimAmount, color = Disposition), stat = "identity")
```


```{r}
TSA_claims %>%
  select(Airport_Code, Airport_Name) %>%
  distinct(Airport_Code)

TSA_claims %>%
  mutate(Incident_Year = as.numeric(format(Incident_Date, "%Y")),
         Claim_Year = as.numeric(format(Date_Received, "%Y"))) %>%
  select(Airport_Code, Incident_Date, Incident_Year, Date_Received, Claim_Year, Claim_Amount, Close_Amount, Disposition) %>%
  group_by(Incident_Year, Claim_Year) %>%
  count(Airport_Code, Incident_Year) %>%
  ggplot() +
  geom_point(mapping = aes(x = n))





```


```{r}

TSA_claims %>%
  select(Date_Received, Incident_Date, Claim_Amount, Close_Amount, Disposition) %>%
  mutate(log_claim = log(Claim_Amount),
         log_close = log(Close_Amount)) %>%
  filter(!is.na(log_claim),
         !is.na(Incident_Date)) %>%
  group_by(Incident_Date, Claim_Amount) %>%
  arrange(Incident_Date) %>%
  ggplot() +
  geom_line(mapping = aes(x = Incident_Date, y = log_claim)) +
  scale_x_date(date_labels = "%b-%y")
   




  

TSA_claims <- TSA_claims %>%
              filter(between(Claim_Amount, quantile(Claim_Amount, 0.1, na.rm = TRUE), x = 
              quantile(Claim_Amount, 0.99, na.rm = TRUE)))
```

Approximately 1600 observations in the dataset have Incident Dates that occus after the date the claim was received (Date_Received). For the purposes of this project, we will modify the Incident_Date of those observations using the duration period derived and adjusting them to precede the Date received
```{r}
TSA_claims %>%
  select(Date_Received, Incident_Date) %>%
  mutate(Duration = Date_Received - Incident_Date) %>%
  mutate(New_Incident_Date = ifelse(Duration < 0, Date_Received + Duration, Incident_Date)) %>%
  mutate(New_Incident_Date = as.Date(New_Incident_Date))
  
  filter(duration < 0) %>%
  mutate(modified_Incident_Date = Date_Received + duration)

  a_age <-  today() - ymd(19790407)
  a_age
  x <- TSA_claims$New_Incident_Date
  
  
  
  
```

```{r}

t <- TSA_claims %>%
  select(Disposition) %>%
  fastDummies::dummy_cols(Disposition)

tplt <- t[ , -1]
tb1 <- table(tplt$`Disposition_Approve in Full`)
tb2 <- table(tplt$Disposition_Settle)
tb3 <- table(tplt$Disposition_Deny)
tb4 <- table(tplt$Disposition_NA)

par(mfrow = c(2,2))

barplot(tb1); barplot(tb2); barplot(tb3); barplot(tb4)

```


```{r}
TSA_claims %>%
  filter(!is.na(Disposition)) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = Disposition), binwidth = 15, stat = "count", color = "slategrey", fill = "slategrey")
                  
```


```{r}

library(dplyr)
library(ggplot2)
#Disposition by Claim Type
TSA_claims %>%
  filter(!is.na(Disposition)) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = Claim_Type), binwidth = 15, stat = "count", color = "slategrey", fill = "slategrey") +
  coord_flip() +
  facet_wrap(Disposition ~.)
```



```{r}
TSA_claims %>%
  filter(Claim_Amount > 10000)
```


```{r}
numeric_to_cat <- function(x, lower, upper, by,
                    sep = "-", above.char = "+") {
  
  labs <- c(paste(seq(lower, upper - by, by = by),
                  seq(lower + by - 1, upper - 1, by = by),
                  sep = sep),
            paste(upper, above.char, sep = ""))
  
  cut(floor(x), breaks = c(seq(lower, upper, by = by), Inf),
      right = FALSE, labels = labs)
}


Claim_bins <- TSA_claims %>%
  select(Claim_Number, Claim_Amount, Close_Amount) %>%
  mutate(Claim_cat = numeric_to_cat(Claim_Amount, 
                                    lower = 0, 
                                    upper = 300000, 
                                    by = 1000),
         Close_cat = numeric_to_cat(Close_Amount, 
                                    lower = 0, 
                                    upper = 250000, 
                                    by = 1000),

  )

Claims <- full_join(TSA_claims, Claim_bins, by = "Claim_Number", suffix = c("", " ")) 

TSA_claims %>%
  select(Claim_Amount, Claim_cat, Close_Amount, Close_cat, Disposition) %>%
  filter(!is.na(Claim_cat),
         !is.na(Close_cat)) %>%
  ggplot() +
  geom_histogram(aes(x = Claim_cat, color = Disposition), binwidth = 2, stat = 'count', color = "lightblue", fill = "lightblue") +
  geom_histogram(aes(x = Close_cat, color = Disposition), binwidth = 2, stat = 'count', color = "grey", fill = "grey") +
  facet_wrap(Disposition ~.) +
  theme_minimal()
 


```



```{r}
library(Hmisc)
library(kableExtra)
library(dplyr)
library(readr)

mean_med <- TSA_claims %>%
  select(Claim_Amount, Close_Amount) %>%
  mutate(mean_ClaimAmount = impute(Claim_Amount, fun = mean),
         mean_CloseAmount = impute(Close_Amount, fun = mean),
         med_ClaimAmount = impute(Claim_Amount, fun = median),
         med_CloseAmount = impute(Close_Amount, fun = median))

View(mean_med)
```



```{r}

im_TSA <- TSA_claims %>%
  select(Claim_Amount, Close_Amount, Disposition)

Disp <- fastDummies::dummy_cols(im_TSA, select_columns = "Disposition", ignore_na = TRUE)
Dispo <- Disp[ , -3]



options(scipen  = 999)

#Dispo <- data.frame(scale(Disp))
model <- glm(Claim_Amount ~ ., data = Dispo)
summary(model)
names(model)
na_Claim <- is.na(im_TSA$Claim_Amount)
im_TSA$Claim_Amount[na_Claim] <- predict(model, newdata = im_TSA[na_Claim, ])
unscale(im_TSA$Claim_Amount[na_Claim])

library(DMwR)
knnImputation(Dispo, k = 10, scale = T, meth = "weighAvg")

library(rpart)
im_tsa <- TSA_claims %>%
  select(Claim_Amount, Close_Amount)
anova_mod <- rpart(Claim_Amount ~ ., data = im_tsa[!is.na(im_tsa$Claim_Amount), ], method="anova", na.action=na.omit)  # since ptratio is numeric.

claim_pred <- predict(anova_mod, im_tsa[is.na(im_tsa$Claim_Amount), ])

actuals <- TSA_claims$Claim_Amount[is.na(im_tsa$Claim_Amount)]
predicteds <- claim_pred
regr.eval(actuals, predicteds)

library(mice)
miceMod <- mice(im_tsa[, !names(im_tsa) %in% "medv"], method="rf")  # perform mice imputation, based on random forests.
miceOutput <- complete(miceMod)  # generate the completed data.
anyNA(miceOutput)

im_tsa <- TSA_claims %>%
  select(Claim_Amount, Close_Amount)

# We will be using mice library in r
library(mice)
# Deterministic regression imputation via mice
imp <- mice(im_tsa, method = "norm.predict", m = 2)

# Store data
data_imp <- complete(imp)
# Multiple Imputation
imp <- mice(im_tsa, m = 5)
#build predictive model
fit <- with(data = imp, lm(y ~ x + z))
#combine results of all 5 models
combine <- pool(fit)
``` 

